---
title: "kornel_kovacs_hw_3"
author: "Kornel Kovacs"
date: '2019 10 01 '
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install the dataset if you don't have it
``install.packages("nycflights13")``

```{r load, eval=TRUE}
library(nycflights13)
flights
View(flights)
library(tidyverse)
```

# Today, we'll cover
- ``filter()``
- ``arrange()``
- ``select()``

# Next week, we'll cover
- ``mutate()``
- ``summarise()``
- ``group_by(), which tells the other verbs to use the data by groups``

All take as first argument a data frame (or tibble) and return a data frame (or tibble).
Together they form the verbs of the tidyverse.


# Filtering (choosing) rows with filter()
```{r choosing, eval=FALSE}
filter(flights, month = 1) # Produces an error
filter(flights, month == 1)
filter(flights, month == 1, day == 1)
filter(flights, dep_time == 517)
```

# dplyr functions don't change the data frame that you give it. They return a new one.
1. Save the filtered data
```{r save, eval=TRUE}
jan1 <- filter(flights, month == 1, day == 1)
jan1
```
2. Assign and print, use (varname <- ...)
```{r assign, eval=TRUE}
(feb1 <- filter(flights, month == 2, day == 1))
```
 
3. Check it really assigned
```{r really, eval=TRUE}
feb1
```

# Some notes on comparisons
```{r notes, eval=TRUE}
sqrt(2)^2 == 2
sqrt(4)^2 == 4
(1/3)*3 == 1
1/49*49 == 1
1/(7^9)*7^9 == 1
```

In short, you can't rely on "It works because it works for what I tried".

# For floating point comparisons, use near() to compare numbers
```{r near, eval=TRUE}
near(sqrt(2)^2, 2)
```

# Multiple constraints |: is 'or' operator
```{r operators, eval=TRUE}
(jan_feb <- filter(flights, month == 1 | month == 2))
(not_jan <- filter(flights, !(month == 1)))
```

# Class exercise: How do we know these actually worked?
```{r justify, eval=TRUE}
filter(not_jan, month == 1)
View(jan_feb)
unique(not_jan$month)
jan <- filter(flights, month == 1)
nrow(flights) == nrow(jan) + nrow(not_jan)

(jan_to_june1 <- filter(flights, month <= 6))
jan_to_june2 <- filter(flights, month %in% c(1,2,3,4,5,6))
```

# Check same number of observations
```{r same_1, eval=TRUE}
nrow(jan_to_june1) == nrow(jan_to_june2)
```

# Class Exercise: What does this do?
```{r mistery, eval=TRUE}
mystery_filter <- filter(flights, !(arr_delay > 120 | dep_delay > 120))
mystery_filter2 <- filter(flights, arr_delay <= 120, dep_delay <= 120)
mystery_filter
mystery_filter2
```

# Vote:
1. All flights that started and landed 120 minutes late
2. All flights that started 120 minutes late or landed 120 minutes late
3. All flights that started less than 120 minutes late or landed less than 120 minutes late
4. All flights that started and landed less than 120 minutes late

# 3. All flights that started less than 120 minutes late or landed less than 120 minutes late
```{r delay, eval=TRUE}
number3 <- filter(flights, arr_delay <= 120 | dep_delay <= 120)
number3 <- filter(flights, arr_delay < 120 | dep_delay < 120)
```

# Class Exercise: get all flights that departed with less than 120 minutes delay, but arrived with more than 120 minutes delay.
```{r flights_graph_3, eval=TRUE}
dep_ok_arr_not <- filter(flights, dep_delay <= 120, arr_delay > 120)

ggplot(data = dep_ok_arr_not,
       mapping = aes(x = dep_delay)) +
  geom_histogram()
```

# Let's look at the data to see what the departure was for planes that arrived 
# late but didn't start quite as late
```{r flights_graph_2, eval=TRUE}
ggplot(data = flights,
       mapping = aes(x = dep_delay)) + 
  geom_histogram()
```

# Filter flights by those that had dep_delay <= 120, then plot histogram
```{r flights_graph, eval=TRUE}
dep_ok <- filter(flights, dep_delay <= 120)
ggplot(data = dep_ok,
       mapping = aes(x = dep_delay)) + 
  geom_histogram()
```

# NA: Not available
```{r na, eval=TRUE}
NA > 5
10 == NA
NA == NA
FALSE & NA
TRUE & NA
NA & FALSE
```

# Let x be Mary's age. We don't know how old she is.
```{r mary, eval=TRUE}
x <- NA
```
# Let y be John's age. We don't know how old he is.
```{r johny, eval=TRUE}
y <- NA
```
# Are John and Mary the same age?
```{r same_2, eval=TRUE}
x == y
```
# We don't know!
```{r nas, eval=TRUE}
NA^0
0 * NA

is.na(x)
df <- tibble(x = c(1, NA, 3))
df
filter(df, x > 1)
filter(df, x > 1 | is.na(x))
```

## arrange()
```{r arrange, eval=TRUE}
flights
arrange(flights, year, month, day)
arrange(flights, dep_delay)
arrange(flights, desc(dep_delay))

arrange(df, x)
arrange(df, desc(x))
```
# Class exercise (do at home): How can we get the missing values at the top?


# Fastest flight
```{r fastest, eval=TRUE}
colnames(flights)
arrange(flights, air_time)
```

# select()
```{r select, eval=TRUE}
select(flights, year, month, day)
select(arrange(flights, air_time), air_time, origin, dest)
```
# That's tedious to write. Hence the pipe.
```{r pipe, eval=TRUE}
flights %>%
  arrange(air_time) %>%
  select(air_time, origin, dest)
```
# Notice that the data doesn't have to be mentioned, and the first argument should not have to be provided
```{r same, eval=TRUE}
select(flights, year:day)
flights %>% select(year:day)
colnames(flights)
```
# dropping cols
```{r drop, eval=TRUE}
select(flights, -(year:day))
```
# Some helper functions
```{r helper, eval=TRUE}
select(flights, starts_with("arr"))
select(flights, -starts_with("arr"))
select(flights, ends_with("hour"))
select(flights, -contains("time"))
```
# Function for renaming columns
```{r rename, eval=TRUE}
rename(flights, destination = dest)
```
# Hard to see if it worked, so...
```{r hard, eval=TRUE}
flights %>% rename(destination = dest) %>% select(year:day, destination)
```
# Moving some columns to the start
```{r moving, eval=TRUE}
select(flights, origin, dest, everything())
```
