---
title: "kornel_kovacs_hw_3"
author: "Kornel Kovacs"
date: '2019 10 01 '
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prerequisites and reads
```{r load, eval=TRUE}
library(tidyverse)
hotels <- read.csv(file = "..//da_data_repo/hotels-vienna/clean/hotels-vienna.csv")
hotels
```

# Overview of data
```{r overview, eval=TRUE}
colnames(hotels)
hotels
```

# Filtering (choosing) rows with filter() and saving.
```{r choosing, eval=TRUE}
(filter(hotels, stars == 4.0))
Alsergrund_4star <- (filter(hotels, stars == 4.0))
```

## Logical operators
### AND - &
```{r and, eval=TRUE}
Alsergrund_4star <- filter(hotels, stars == 4.0 & neighbourhood == 'Alsergrund')
Alsergrund_4star <- filter(hotels, stars == 4.0,  neighbourhood == 'Alsergrund')
```

### OR - |

```{r or, eval=TRUE}
Alsergrund_or_4star <- filter(hotels, stars == 4.0 | neighbourhood == 'Alsergrund')
```

### NOT - !

```{r not, eval=TRUE}
not_4star_in_Alsergrund_4star <- filter(hotels, stars != 4.0)
```

### Let's combine them!

```{r combine, eval=TRUE}
not_4star_in_Alsergrund <- filter(hotels, stars != 4.0 & neighbourhood == 'Alsergrund')
```

### Trick

```{r trick, eval=TRUE}
four_five_stars <- filter(hotels, stars %in% c(4.0,5.0))
```

### Do not overcomplicate, do not use negation unless necessary!

```{r overcomplicate, eval=TRUE}
filter(hotels, !(rating > 4.0 | stars > 4.0))
filter(hotels, rating <= 4.0, stars <= 4.0)
```

## I skipped the NA part, because it would have been copy-pasting, which I try to avoid.

## arrange()

### Worst hotels according to ratings
### Note that in a real analysis we should take into account the number of ratings as well.
```{r worst, eval=TRUE}
arrange(hotels, rating)
```

### Best hotels according to ratings
```{r best, eval=TRUE}
arrange(hotels, desc(rating))
```
