---
title: "kornel_kovacs_hw_3_hotels"
author: "Kornel Kovacs"
date: '2019 10 01 '
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Prerequisites and reads
```{r load, eval=TRUE}
library(tidyverse)
hotels <- read.csv(file = "..//da_data_repo/hotels-vienna/clean/hotels-vienna.csv")
head(hotels)
```

# 2. Overview of data
```{r overview, eval=TRUE}
colnames(hotels)
```

# 3. Filtering (choosing) rows with filter() and saving
```{r choosing, eval=TRUE}
head((filter(hotels, stars == 4.0)))
Alsergrund_4star <- (filter(hotels, stars == 4.0))
```

## 3.1 Logical operators
### 3.1.1 AND - &
```{r and, eval=TRUE}
Alsergrund_4star <- filter(hotels, stars == 4.0 & neighbourhood == 'Alsergrund')
Alsergrund_4star <- filter(hotels, stars == 4.0,  neighbourhood == 'Alsergrund')
```

### 3.1.2 OR - |

```{r or, eval=TRUE}
Alsergrund_or_4star <- filter(hotels, stars == 4.0 | neighbourhood == 'Alsergrund')
```

### 3.1.3 NOT - !

```{r not, eval=TRUE}
not_4star_in_Alsergrund_4star <- filter(hotels, stars != 4.0)
```

### 3.1.4 Let's combine them!

```{r combine, eval=TRUE}
not_4star_in_Alsergrund <- filter(hotels, stars != 4.0 & neighbourhood == 'Alsergrund')
```

### 3.1.5 Trick

```{r trick, eval=TRUE}
four_five_stars <- filter(hotels, stars %in% c(4.0,5.0))
```

### 3.1.6 Do not overcomplicate, do not use negation unless necessary!

```{r overcomplicate, eval=TRUE}
head(filter(hotels, !(rating > 4.0 | stars > 4.0)))
head(filter(hotels, rating <= 4.0, stars <= 4.0))
```

## 4. I skipped the NA part, because it would have been copy-pasting, which I try to avoid.

## 5. arrange()

### 5.1 Worst hotels according to ratings
Note that in a real analysis we should take into account the number of ratings as well.
```{r worst, eval=TRUE}
head(arrange(hotels, rating))
```

### 5.2 Best hotels according to ratings
```{r best, eval=TRUE}
head(arrange(hotels, desc(rating)))
```

## 5. select()

```{r select_1, eval=TRUE}
head(select(hotels, accommodation_type , nnights , rating))
```

```{r select_2, eval=TRUE}
head(select(hotels, accommodation_type:rating))
```

```{r select_3, eval=TRUE}
head(select(hotels, -(center1label:rating)))
```

```{r select_4, eval=TRUE}
head(select(hotels, accommodation_type , rating, everything()))
```

```{r rename, eval=TRUE}
head(rename(hotels, customer_rating = rating))
```


