---
title: "Assignment 12"
author: "Kornel Kovacs"
date: '2019 11 22 '
output:
  html_document: default
  pdf_document: default
---

## General setup for the whole project.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pbapply)
library(plotrix)
```

```{r}
df <- read_csv("./hotels-vienna.csv")
```

## Exercise 1

I check the mean of the price in the data.
```{r}
mean(df$price)
```

I am simulating 1 000 samples with replacement and saving their means in a df.
```{r}
bootstrap_df <- data_frame(num = 1:1000) %>% 
    group_by(num) %>% 
    mutate(means = mean(sample(df$price, replace = TRUE)))
```

Therefore, the bootstrap standard deviation of this mean (aka ‘the standard error’) can be calculated easily.

```{r}
sd(bootstrap_df$means)
```

## Exercise 2

I designed the function as the exercise wished.

```{r}
bootstrap_mean <- function(B,v) {
  bootstrap_df <- data_frame(num = 1:B) %>% 
    group_by(num) %>% 
    mutate(means = mean(sample(v, replace = TRUE)))
  return(sd(bootstrap_df$means))
}
```

I apply the function on a bunch of values for the plot.

```{r}
simulated_deviations <- unlist(pblapply(2:1000, bootstrap_mean, v=df$price))
df_to_plot <- data.frame(number_of_simulations = 2:1000, 
                         standard_dev = simulated_deviations)
```

```{r}
ggplot(df_to_plot, aes(x = number_of_simulations, y = standard_dev)) + 
  geom_point() +
  xlab("Number of simulations") +
  ylab("Standard deviation of mean")
```

## Exercise 3

I found a nice package called ``plotrix`` which has a function that calculates the standard error for ``price``.

```{r}
std.error(df$price)
```

## Exercise 4

I define the function to calculate standard deviation for median values.

```{r}
bootstrap_median <- function(B,v) {
  bootstrap_df <- data_frame(num = 1:B) %>% 
    group_by(num) %>% 
    mutate(medians = median(sample(v, replace = TRUE)))
  return(sd(bootstrap_df$medians))
}
```

I also try it out.

```{r}
bootstrap_median(1000, df$price)
```

## Exercise 5

I defined my Swiss army knife, a multifunctional method.

```{r}
bootstrap_func <- function(B,v, method) {
  bootstrap_df <- data_frame(num = 1:B) %>% 
    group_by(num) %>% 
    mutate(metric_value = method(sample(v, replace = TRUE)))
  return(sd(bootstrap_df$metric_value))
}
```

Example calls:

```{r}
bootstrap_func(1000, df$price, sd)
```

```{r}
bootstrap_func(1000, df$price, max)
```

```{r}
bootstrap_func(1000, df$price, mean)
```

```{r}
bootstrap_func(1000, df$price, median)
```

As we do not have a basic function for the top quartile, some hacking is needed for this case.

```{r}
bootstrap_func(1000, df$price,function(metric) quantile(metric)[4])
```

## Exercise 6

```{r}
get_coeff <- function(data) {
  ide még kell a modell
  aztán kiszedni a koefficienseket
  
}
lm( df$price ~ df$ratingta)
vmi$coefficients[2]
```


























