---
title: "Assignment 8"
author: "Kornel Kovacs"
date: "10/31/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library('haven')
```


# Exercise 1:
I created a project for this assignment in a folder.

# Exercise 2:
All done

# Exercise 3:
Reading in from the CSV:
```{r}
df_csv <- read_csv('./data_repo/hotels-europe/clean/hotels-europe_price.csv')
```

Reading in from the DTA:
```{r}
df_dta <- read_dta('./data_repo/hotels-europe/clean/hotels-europe_price.dta')
```

Let us see whether they have the same lenght or not. Spoiler: They do.
```{r}
nrow(df_csv)
nrow(df_dta)
```

Let us see whether they have the same head and tail or not. Spoiler: They do.
```{r}
head(df_csv)
head(df_dta)
```

```{r}
tail(df_csv)
tail(df_dta)
```

I was wondering whether we have the same columns and types of them are consistent or not. Spoiler: They do.
```{r}
sapply(df_csv, class)
sapply(df_dta, class)
```

Finally, there are more sophisticated ways of showing whether they are identical or not. I should have started with this one, but I wanted to show several aspects of testing before coming up with these functions.

All values are the same or not:
```{r, warning=FALSE}
all.equal(df_csv, df_dta)
```

Okay..but these are only the values. Are they dentical objects as well?
```{r}
identical(df_dta, df_csv)
```

**Somehow, they are not identical according to the function ``identical()``**

# Exercise 4:
Creating a sample of 200 to write out.
```{r}
df_csv_200 <- df_csv[sample(nrow(df_csv), 200), ]
```

Actually writing it out. Notice, I already did the alteration part as ``write_csv2()`` uses ``;`` as a separator.
```{r}
write_csv2(df_csv_200, "hotels_cleaned_200.csv")
```

Now, try reading it back in.
```{r, warning=FALSE}
df_csv_200 <- read_csv("hotels_cleaned_200.csv")
head(df_csv_200)
```

We do not see a structure, just the whole thing in one column. The problem is that ``read_csv()`` is looking for ``,`` to separate values by, but it doesn't find any of such a kind. As a solution we either have to specify the value of the separator or use another function which has ``;`` as the default separator value. ``read_csv2()`` is such a function. I chose this way of screwing up the dataset because it is very often the case that we have something else as a separator other than ``,``.

# Exercise 5:

We have data from WHO to make it tidy.
```{r}
head(who)
```

We use the function ``gather()`` to get rid of several columns leaving out NAs.
```{r}
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
head(who1)
```

We can count the keys we just created to get a sense we have done it right.
```{r}
who1 %>% 
  count(key)
```

Correcting some inconsistency concerning the values of the columns ``key``.
```{r}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
head(who2)
```

We split values and create new columns.
```{r}
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
head(who3)
```

We drop columns we do not use or redundant.
```{r}
who4 <- who3 %>% 
  select(-new, -iso2, -iso3)
head(who4)
```

You may have noticed that ``sexage`` is still not a standalone columns. Let's split that!
```{r}
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
head(who5)
```

Putting it all together in a huge pipe.
```{r}
who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

Now, we have a nice tidy table. We can start the analysis!

# Exercise 6:

I am supposed to tidy the following dataset.
```{r, data_exercise_t}
df <- tibble(name = c("A123", "B456"), 
             age = c(30, 60), 
             answer1 = c(0, 1), 
             answer2 = c(1,1), 
             answer3 = c(1,0), 
             answer4 = c(0,0))
head(df)
```

First, I realize that they are the ``answerN`` columns that should be modified. If I want to add more answer types(?) as dimensions to this table I always have to create a new column. I do not want that. I would only want to add a new row to the existing table. Accordingly, I keep ``name`` and ``age`` and add a column to assess the 'type' of the answer and the value of it.

Make it tidy with ``gather()``:
```{r}
df_tidy <- df %>%
  gather(`answer1`, 
         `answer2`, 
         `answer3`, 
         `answer4`, 
         key = "type of answer", 
         value = "value of the answer")
head(df_tidy)
```
